### 2014-06-06 scrap.tex
coarse grid run

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{gensymb}
\usepackage{amsmath}
\usepackage{color}
\usepackage{float}

\begin{document}
The whole dispersion relation:\\
$B_{0}^{4} C^{2} k_{x}^{4} k_{z}^{4} + 2 B_{0}^{4} C^{2} k_{x}^{2} k_{z}^{6} + B_{0}^{4} C^{2} k_{z}^{8} - 2 B_{0}^{2} C \Omega^{2} k_{x}^{2} k_{z}^{4} q - 2 B_{0}^{2} C \Omega^{2} k_{z}^{6} q - \frac{B_{0}^{2} C \Omega k_{x}^{4} k_{z}^{4} q}{Rm} + 2 \frac{B_{0}^{2} C \Omega k_{x}^{4} k_{z}^{4}}{Rm} + \frac{B_{0}^{2} C \Omega k_{x}^{3} k_{z}^{5} q}{Rm} - 2 \frac{B_{0}^{2} C \Omega k_{x}^{3} k_{z}^{5}}{Rm} - \frac{B_{0}^{2} C \Omega k_{x}^{2} k_{z}^{6} q}{Rm} + 2 \frac{B_{0}^{2} C \Omega k_{x}^{2} k_{z}^{6}}{Rm} + \frac{B_{0}^{2} C \Omega k_{x} k_{z}^{7} q}{Rm} - 2 \frac{B_{0}^{2} C \Omega k_{x} k_{z}^{7}}{Rm} + \frac{B_{0}^{2} C \Omega k_{x}^{4} k_{z}^{4} q}{R} - \frac{B_{0}^{2} C \Omega k_{x}^{3} k_{z}^{5} q}{R} + \frac{B_{0}^{2} C \Omega k_{x}^{2} k_{z}^{6} q}{R} - \frac{B_{0}^{2} C \Omega k_{x} k_{z}^{7} q}{R} + 2 B_{0}^{2} C k_{x}^{4} k_{z}^{2} s^{2} + 4 B_{0}^{2} C k_{x}^{2} k_{z}^{4} s^{2} + 2 B_{0}^{2} C k_{z}^{6} s^{2} - \frac{B_{0}^{2} C k_{x}^{5} k_{z}^{3} s}{Rm} + 2 \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{4} s}{Rm} - 2 \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{5} s}{Rm} + 2 \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{6} s}{Rm} - \frac{B_{0}^{2} C k_{x} k_{z}^{7} s}{Rm} - \frac{B_{0}^{2} C k_{x}^{5} k_{z}^{3} s}{R} + 2 \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{4} s}{R} - 2 \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{5} s}{R} + 2 \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{6} s}{R} - \frac{B_{0}^{2} C k_{x} k_{z}^{7} s}{R} + \frac{B_{0}^{2} C k_{x}^{6} k_{z}^{4}}{R Rm} - 4 \frac{B_{0}^{2} C k_{x}^{5} k_{z}^{5}}{R Rm} + 6 \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{6}}{R Rm} - 4 \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{7}}{R Rm} + \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{8}}{R Rm} + 2 \frac{\Omega^{3} k_{x}^{2} k_{z}^{4} q^{2}}{Rm} - 4 \frac{\Omega^{3} k_{x}^{2} k_{z}^{4} q}{Rm} - 2 \frac{\Omega^{3} k_{x} k_{z}^{5} q^{2}}{Rm} + 4 \frac{\Omega^{3} k_{x} k_{z}^{5} q}{Rm} - 2 \Omega^{2} k_{x}^{2} k_{z}^{2} q s^{2} + 4 \Omega^{2} k_{x}^{2} k_{z}^{2} s^{2} - 2 \Omega^{2} k_{z}^{4} q s^{2} + 4 \Omega^{2} k_{z}^{4} s^{2} + 2 \frac{\Omega^{2} k_{x}^{3} k_{z}^{3} q s}{Rm} - 4 \frac{\Omega^{2} k_{x}^{3} k_{z}^{3} s}{Rm} - 4 \frac{\Omega^{2} k_{x}^{2} k_{z}^{4} q s}{Rm} + 8 \frac{\Omega^{2} k_{x}^{2} k_{z}^{4} s}{Rm} + 2 \frac{\Omega^{2} k_{x} k_{z}^{5} q s}{Rm} - 4 \frac{\Omega^{2} k_{x} k_{z}^{5} s}{Rm} - \frac{\Omega^{2} k_{x}^{4} k_{z}^{4} q^{2}}{R Rm} + 2 \frac{\Omega^{2} k_{x}^{4} k_{z}^{4} q}{R Rm} + 2 \frac{\Omega^{2} k_{x}^{3} k_{z}^{5} q^{2}}{R Rm} - 4 \frac{\Omega^{2} k_{x}^{3} k_{z}^{5} q}{R Rm} - \frac{\Omega^{2} k_{x}^{2} k_{z}^{6} q^{2}}{R Rm} + 2 \frac{\Omega^{2} k_{x}^{2} k_{z}^{6} q}{R Rm} - \frac{\Omega k_{x}^{4} k_{z}^{2} q s^{2}}{Rm} + \frac{\Omega k_{x}^{3} k_{z}^{3} q s^{2}}{Rm} - \frac{\Omega k_{x}^{2} k_{z}^{4} q s^{2}}{Rm} + \frac{\Omega k_{x} k_{z}^{5} q s^{2}}{Rm} + \frac{\Omega k_{x}^{4} k_{z}^{2} q s^{2}}{R} - 2 \frac{\Omega k_{x}^{4} k_{z}^{2} s^{2}}{R} - \frac{\Omega k_{x}^{3} k_{z}^{3} q s^{2}}{R} + 2 \frac{\Omega k_{x}^{3} k_{z}^{3} s^{2}}{R} + \frac{\Omega k_{x}^{2} k_{z}^{4} q s^{2}}{R} - 2 \frac{\Omega k_{x}^{2} k_{z}^{4} s^{2}}{R} - \frac{\Omega k_{x} k_{z}^{5} q s^{2}}{R} + 2 \frac{\Omega k_{x} k_{z}^{5} s^{2}}{R} + 2 \frac{\Omega k_{x}^{5} k_{z}^{3} s}{R Rm} - 6 \frac{\Omega k_{x}^{4} k_{z}^{4} s}{R Rm} + 6 \frac{\Omega k_{x}^{3} k_{z}^{5} s}{R Rm} - 2 \frac{\Omega k_{x}^{2} k_{z}^{6} s}{R Rm} + k_{x}^{4} s^{4} + 2 k_{x}^{2} k_{z}^{2} s^{4} + k_{z}^{4} s^{4} - \frac{k_{x}^{5} k_{z} s^{3}}{Rm} + 2 \frac{k_{x}^{4} k_{z}^{2} s^{3}}{Rm} - 2 \frac{k_{x}^{3} k_{z}^{3} s^{3}}{Rm} + 2 \frac{k_{x}^{2} k_{z}^{4} s^{3}}{Rm} - \frac{k_{x} k_{z}^{5} s^{3}}{Rm} - \frac{k_{x}^{5} k_{z} s^{3}}{R} + 2 \frac{k_{x}^{4} k_{z}^{2} s^{3}}{R} - 2 \frac{k_{x}^{3} k_{z}^{3} s^{3}}{R} + 2 \frac{k_{x}^{2} k_{z}^{4} s^{3}}{R} - \frac{k_{x} k_{z}^{5} s^{3}}{R} + \frac{k_{x}^{6} k_{z}^{2} s^{2}}{R Rm} - 4 \frac{k_{x}^{5} k_{z}^{3} s^{2}}{R Rm} + 6 \frac{k_{x}^{4} k_{z}^{4} s^{2}}{R Rm} - 4 \frac{k_{x}^{3} k_{z}^{5} s^{2}}{R Rm} + \frac{k_{x}^{2} k_{z}^{6} s^{2}}{R Rm}$ \\

Dispersion relation minus a4: are there any 0-order s terms?\\
$B_{0}^{4} C^{2} k_{x}^{4} k_{z}^{4} + 2 B_{0}^{4} C^{2} k_{x}^{2} k_{z}^{6} + B_{0}^{4} C^{2} k_{z}^{8} + 2 B_{0}^{2} C k_{x}^{4} k_{z}^{2} s^{2} - 2 B_{0}^{2} C k_{x}^{2} k_{z}^{4} \omega^{2} q + 4 B_{0}^{2} C k_{x}^{2} k_{z}^{4} s^{2} - 2 B_{0}^{2} C k_{z}^{6} \omega^{2} q + 2 B_{0}^{2} C k_{z}^{6} s^{2} - \frac{B_{0}^{2} C k_{x}^{5} k_{z}^{3} s}{Rm} - \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{4} \omega q}{Rm} + 2 \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{4} \omega}{Rm} + 2 \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{4} s}{Rm} + \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{5} \omega q}{Rm} - 2 \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{5} \omega}{Rm} - 2 \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{5} s}{Rm} - \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{6} \omega q}{Rm} + 2 \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{6} \omega}{Rm} + 2 \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{6} s}{Rm} + \frac{B_{0}^{2} C k_{x} k_{z}^{7} \omega q}{Rm} - 2 \frac{B_{0}^{2} C k_{x} k_{z}^{7} \omega}{Rm} - \frac{B_{0}^{2} C k_{x} k_{z}^{7} s}{Rm} - \frac{B_{0}^{2} C k_{x}^{5} k_{z}^{3} s}{R} + \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{4} \omega q}{R} + 2 \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{4} s}{R} - \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{5} \omega q}{R} - 2 \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{5} s}{R} + \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{6} \omega q}{R} + 2 \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{6} s}{R} - \frac{B_{0}^{2} C k_{x} k_{z}^{7} \omega q}{R} - \frac{B_{0}^{2} C k_{x} k_{z}^{7} s}{R} + \frac{B_{0}^{2} C k_{x}^{6} k_{z}^{4}}{R Rm} - 4 \frac{B_{0}^{2} C k_{x}^{5} k_{z}^{5}}{R Rm} + 6 \frac{B_{0}^{2} C k_{x}^{4} k_{z}^{6}}{R Rm} - 4 \frac{B_{0}^{2} C k_{x}^{3} k_{z}^{7}}{R Rm} + \frac{B_{0}^{2} C k_{x}^{2} k_{z}^{8}}{R Rm} + k_{x}^{4} s^{4} - 2 k_{x}^{2} k_{z}^{2} \omega^{2} q s^{2} + 4 k_{x}^{2} k_{z}^{2} \omega^{2} s^{2} + 2 k_{x}^{2} k_{z}^{2} s^{4} - 2 k_{z}^{4} \omega^{2} q s^{2} + 4 k_{z}^{4} \omega^{2} s^{2} + k_{z}^{4} s^{4} - \frac{k_{x}^{5} k_{z} s^{3}}{Rm} - \frac{k_{x}^{4} k_{z}^{2} \omega q s^{2}}{Rm} + 2 \frac{k_{x}^{4} k_{z}^{2} s^{3}}{Rm} + 2 \frac{k_{x}^{3} k_{z}^{3} \omega^{2} q s}{Rm} - 4 \frac{k_{x}^{3} k_{z}^{3} \omega^{2} s}{Rm} + \frac{k_{x}^{3} k_{z}^{3} \omega q s^{2}}{Rm} - 2 \frac{k_{x}^{3} k_{z}^{3} s^{3}}{Rm} + 2 \frac{k_{x}^{2} k_{z}^{4} \omega^{3} q^{2}}{Rm} - 4 \frac{k_{x}^{2} k_{z}^{4} \omega^{3} q}{Rm} - 4 \frac{k_{x}^{2} k_{z}^{4} \omega^{2} q s}{Rm} + 8 \frac{k_{x}^{2} k_{z}^{4} \omega^{2} s}{Rm} - \frac{k_{x}^{2} k_{z}^{4} \omega q s^{2}}{Rm} + 2 \frac{k_{x}^{2} k_{z}^{4} s^{3}}{Rm} - 2 \frac{k_{x} k_{z}^{5} \omega^{3} q^{2}}{Rm} + 4 \frac{k_{x} k_{z}^{5} \omega^{3} q}{Rm} + 2 \frac{k_{x} k_{z}^{5} \omega^{2} q s}{Rm} - 4 \frac{k_{x} k_{z}^{5} \omega^{2} s}{Rm} + \frac{k_{x} k_{z}^{5} \omega q s^{2}}{Rm} - \frac{k_{x} k_{z}^{5} s^{3}}{Rm} - \frac{k_{x}^{5} k_{z} s^{3}}{R} + \frac{k_{x}^{4} k_{z}^{2} \omega q s^{2}}{R} - 2 \frac{k_{x}^{4} k_{z}^{2} \omega s^{2}}{R} + 2 \frac{k_{x}^{4} k_{z}^{2} s^{3}}{R} - \frac{k_{x}^{3} k_{z}^{3} \omega q s^{2}}{R} + 2 \frac{k_{x}^{3} k_{z}^{3} \omega s^{2}}{R} - 2 \frac{k_{x}^{3} k_{z}^{3} s^{3}}{R} + \frac{k_{x}^{2} k_{z}^{4} \omega q s^{2}}{R} - 2 \frac{k_{x}^{2} k_{z}^{4} \omega s^{2}}{R} + 2 \frac{k_{x}^{2} k_{z}^{4} s^{3}}{R} - \frac{k_{x} k_{z}^{5} \omega q s^{2}}{R} + 2 \frac{k_{x} k_{z}^{5} \omega s^{2}}{R} - \frac{k_{x} k_{z}^{5} s^{3}}{R} + \frac{k_{x}^{6} k_{z}^{2} s^{2}}{R Rm} + 2 \frac{k_{x}^{5} k_{z}^{3} \omega s}{R Rm} - 4 \frac{k_{x}^{5} k_{z}^{3} s^{2}}{R Rm} - \frac{k_{x}^{4} k_{z}^{4} \omega^{2} q^{2}}{R Rm} + 2 \frac{k_{x}^{4} k_{z}^{4} \omega^{2} q}{R Rm} - 6 \frac{k_{x}^{4} k_{z}^{4} \omega s}{R Rm} + 6 \frac{k_{x}^{4} k_{z}^{4} s^{2}}{R Rm} + 2 \frac{k_{x}^{3} k_{z}^{5} \omega^{2} q^{2}}{R Rm} - 4 \frac{k_{x}^{3} k_{z}^{5} \omega^{2} q}{R Rm} + 6 \frac{k_{x}^{3} k_{z}^{5} \omega s}{R Rm} - 4 \frac{k_{x}^{3} k_{z}^{5} s^{2}}{R Rm} - \frac{k_{x}^{2} k_{z}^{6} \omega^{2} q^{2}}{R Rm} + 2 \frac{k_{x}^{2} k_{z}^{6} \omega^{2} q}{R Rm} - 2 \frac{k_{x}^{2} k_{z}^{6} \omega s}{R Rm} + \frac{k_{x}^{2} k_{z}^{6} s^{2}}{R Rm} - \frac{- 2 C^{4} Rm^{4} k_{z}^{4} q + C^{3} Rm^{2} k_{z}^{2} \left(k_{x}^{2} + k_{z}^{2}\right)^{2} \left(- 2 q + 4\right) + C \left(k_{x}^{2} + k_{z}^{2}\right) \left(C^{2} Rm^{2} k_{z}^{2} + \frac{C Rm \left(k_{x}^{2} + k_{z}^{2}\right)^{2}}{R}\right)^{2}}{C^{3} Rm^{4}}$

\end{document}