### 2016-04-12 paper/wide_gap_notes.tex
math notes url changed

\documentclass{paper}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\hyd}    {{\rm H}}
\newcommand{\vobs}{v_{\rm obs}}
\newcommand{\hi}{H{\sc i}~}
\newcommand{\hii}{H{\sc ii}~}
\newcommand{\hia}{H{\sc i}}
\newcommand{\hiia}{H{\sc ii}}

\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{gensymb}
\usepackage{amsmath}
\usepackage{color}
%\usepackage{appendix}
\usepackage[titletoc,title]{appendix}
\usepackage{cleveref}
\usepackage{mathtools}

\makeatletter
\setlength{\@fptop}{0pt}
\setlength{\@fpbot}{0pt}
\setlength{\@fpsep}{0pt}
\makeatother

\newcommand{\citei}[1]{\citeauthor{#1} \citeyear{#1}}
\newcommand{\unit}[1]{\textrm{ #1}}
\newcommand{\kms}{km ${\rm s^{-1}}$}
\newcommand{\kmsa}{km ${\rm s^{-1}}$}
\newcommand{\citeia}[2]{\citeauthor{#1}~(\citeyear{#1};~#2)}

\newcommand{\uphi}{\ensuremath{u_\phi}}
\newcommand{\rhat}{\ensuremath{\mathbf{\hat{r}}}}
\newcommand{\phihat}{\ensuremath{\mathbf{\hat{\phi}}}}
\newcommand{\zhat}{\ensuremath{\mathbf{\hat{z}}}}

\newcommand\reye{\mathrm{Re}}
\newcommand\reym{\mathrm{Rm}}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\begin{document}

\title{Notes on Wide Gap}

\section{Basic Equations}
\label{sec:equations}
The so-called Stokes stream function, used in axisymmetric situations, is given by 
\begin{equation}
  \label{eq:stokes}
  \mathbf{u} \, = \, \left[\begin{matrix}
\frac{1}{r} \partial_z \psi\ \rhat\\
\uphi \ \phihat\\
-\frac{1}{r} \partial_r \psi\ \zhat\\
\end{matrix}\right];
\end{equation}
here we define $A$ in the same way. \\

Using the definitions in

\begin{multline}
  \label{eq:psi}
\partial_t \left[ \frac{1}{r} \left(\nabla^2 \psi - \frac{2 \partial_r \psi}{r} \right) \right] + \frac{1}{r^2} J(\psi, \nabla^2 A - \frac{2 \partial_r \psi}{r}) = \frac{\partial_z A}{r^3} \left( \nabla^2 A - \frac{2 \partial_r A}{r}\right) \\
+ \frac{1}{r} J\left(A,\frac{1}{r}\left( \nabla^2 A - \frac{2 \partial_r A}{r}\right)\right) - \frac{2 B_\phi \partial_z B_\phi}{r}\\
+ \nu \left\{ \nabla^2 \left[ \frac{1}{r} \left(\nabla^2 \psi - \frac{2 \partial_r \psi}{r}\right) \right]  -\frac{1}{r^2} \left(\nabla^2 \psi - \frac{2 \partial_r \psi}{r}\right) \right\}
\end{multline}

For the expanded form of the $\Psi$ equation, Susan gets:

\begin{equation}
  \label{eq:uy}
\partial_t \uphi + \frac{J(\psi, \uphi)}{r} + \frac{\uphi \partial_z \psi}{r^2} = \frac{J(A, B_\phi)}{r} + \frac{B_\phi \partial_z A}{r^2} + \nu \left( \nabla^2 \uphi - \frac{\uphi}{r}\right)
\end{equation}

%\begin{equation}
 % \label{eq:A}
%\partial_t A = \frac{1}{r} J(A, \psi) + \eta \left( \nabla^2 A - \frac{2 \partial_r A}{r} \right)
%\end{equation}

%Susan confirms correct (just switched eta -> 1/Rm)
\beq
 \label{eq:A}
\partial_t A = \frac{1}{r} J(A, \psi) + \frac{1}{\reym} \left( \nabla^2 A - \frac{2 \partial_r A}{r} \right)
%\partial_t A = \frac{1}{r} J(A, \Psi) + \frac{1}{Rm} \left[ -\frac{1}{r} \partial_r A + \partial_r^2 A + \partial_z^2 A\right]
\eeq
%Susan got:
%beq
%\partial_t A = \frac{2}{r} J(A, \Psi) + \frac{1}{Rm} \left[ \nabla^2 A - \frac{2}{r} \partial_r A \right]
%\eeq
%....will check.

\begin{multline}
  \label{eq:By} % Susan confirms correct 1/14/16
\partial_t B_\phi = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi)\\
+ \frac{1}{r^2} B_\phi \partial_z \psi - \frac{1}{r^2} \uphi \partial_z A + \eta \left(\nabla^2 B_\phi - \frac{1}{r^2} B_\phi \right)
\end{multline}

\section{Detailed Derivation of $\Psi$ Equation}
The $\Psi$ equation, governing the x- and z-components of the velocity, is particularly tricky to derive so I will write out the steps here.

1. Find $\rhat$ and $\zhat$ components of the momentum equation, i.e.:

\beq
\partial_t u_z + \left[u \cdot \nabla u \right]_z = \left[ \left( \nabla \times B \right) \times B \right]_z + \frac{1}{\reye}\left[ \nabla^2 u \right] _z
\eeq

We sub in our stream/flux function notation and expand the operators in cylindrical coordinates. Then take $\partial_r$ of the resulting equation to obtain:

\begin{multline}
\frac{1}{r^2} \partial_t \partial_r \Psi - \frac{1}{r} \partial_t \partial_r^2 \Psi - \frac{3}{r^4}\partial_z \Psi \partial_r \Psi + \frac{1}{r^3} \partial_r \left( \partial_z \Psi \partial_r \Psi \right) + \frac{2}{r^3} \partial_z \Psi \partial_r^2 \Psi - \frac{1}{r^2} \partial_r \left(\partial_z \Psi \partial_r^2 \Psi \right)\\
- \frac{2}{r^3} \partial_r \Psi \partial_r\partial_z \Psi + \frac{1}{r^2} \partial_r \left(\partial_r \Psi \partial_r \partial_z \Psi\right) = \\
\partial_r\left(B_\phi \partial_z B_\phi\right) + \frac{2}{r^3} \partial_z^2 A \partial_z A - \frac{1}{r^2} \partial_r \left(\partial_z^2 A \partial_z A\right) + \frac{3}{r^4} \partial_z A \partial_r A - \frac{1}{r^3} \partial_r \left(\partial_z A \partial_r A \right) - \frac{2}{r^3} \partial_z A \partial_r^2 A \\
+ \frac{1}{r^2} \partial_r \left(\partial_z A \partial_r^2 A \right) 
+ \frac{1}{\reye} \left[\frac{3}{r^4} \partial_r \Psi - \frac{3}{r^3} \partial_r^2 \Psi + \frac{2}{r^2} \partial_r^3 \Psi - \frac{1}{r} \partial_r^4 \Psi\right]
\end{multline}

Repeat this process for the $\rhat$ component of the momentum equation,

\beq
\partial_t u_r + \left[u \cdot \nabla u \right]_r = \left[ \left( \nabla \times B \right) \times B \right]_r + \frac{1}{\reye}\left[ \nabla^2 u \right] _r
\eeq

and take $\partial_z$ of the expanded equation to obtain

\begin{multline}
\frac{1}{r} \partial_t \partial_z^2 \Psi - \frac{1}{r^3} \partial_z \left(\partial_z \Psi \partial_z \Psi\right) + \frac{1}{r^2} \partial_z\left(\partial_z\Psi \partial_z\partial_r \Psi\right) - \frac{1}{r^2} \partial_z \left(\partial_r \Psi \partial_z^2 \Psi\right) - \frac{1}{r} 2 u_\phi \partial_z u_\phi \\
= - \frac{1}{r^2} \partial_z^3 A \partial_r A - \frac{1}{r^2} \partial_z^2 A \partial_r \partial_z A + \frac{2}{r^3} \partial_r\partial_z A \partial_r A - \frac{1}{r^2} \partial_r^2 \partial_z A \partial_rA - \frac{1}{r^2} \partial_r^2 A \partial_r \partial_z A \\
+ \frac{1}{\reye} \left[ -\frac{1}{r^2} \partial_z^2 \partial_r \Psi + \frac{1}{r} \partial_z^2 \partial_r^2 \Psi + \frac{1}{r} \partial_z^4\Psi \right]
\end{multline}

It is clear from the $\partial_t$ terms that we must combine these equations by subtracting the $\zhat$ equation from the $\rhat$ equation.

When we do, we can simplify the LHS of the equation to:

\beq
\frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) + J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) - \frac{1}{r} 2 u_\phi \partial_z u_\phi
\eeq

Note that the relevant quantity appears to be $\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi$, and that the $\frac{1}{r^2}$ in the second term cannot come out of the Jacobian (a point of disagreement with Jeff's equation above). Also I'm confused why Jeff's has no $u_\phi$ term. The RHS of this equation is significantly more complicated.

RHS viscous term:

\beq
\frac{1}{\reye} \left[ \nabla^2 \left(\frac{1}{r} \nabla^2 \Psi\right) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right]
\eeq

Full $\Psi$ equation according to Susan:

\begin{multline}
\frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) + J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) - \frac{1}{r} 2 u_\phi \partial_z u_\phi \\
= J\left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_rA\right) \right) - \frac{2}{r}B_\phi \partial_z B_\phi \\
+ \frac{1}{\reye} \left[ \nabla^2 \left(\frac{1}{r} \nabla^2 \Psi\right) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right]
\end{multline}

Note that this is actually beautifully symmetric. Except the viscous term which still seems clunky.....

The derivation of the non-viscous term on the righthand side of the momentum equation ($\mathbf{J} \times \mathbf{B}$) is as follows.

\beq
\partial_z\left(\left[\left(\nabla \times B \right) \times B \right]_r\right) - \partial_r \left(\left[\left(\nabla \times B\right) \times B \right]_z\right)
\eeq

\beq
= \partial_z\left(\left[\left(\partial_z B_r - \partial_r B_z \right) B_z - \left(\frac{1}{r}\partial_r\left(r B_\phi\right)\right) B_\phi \right] \right) - \partial_r \left(\left[\left(-\partial_z B_\phi\right)B_\phi - \left(\partial_z B_r - \partial_r B_z \right) B_r\right]\right)
\eeq

\begin{multline}
= -\frac{1}{r^2} \partial_z^3 A \partial_r A + \frac{1}{r^3} \partial_r \partial_z A \partial_r A - \frac{1}{r^2} \partial_r^2 \partial_z A \partial_r A - \frac{2}{r^3} \partial_z^2 A \partial_z A \\
+ \frac{1}{r^2} \partial_z^2 \partial_r A \partial_z A + \frac{3}{r^4} \partial_r A \partial_z A - \frac{3}{r^3} \partial_r^2 A \partial_z A + \frac{1}{r^2} \partial_r^3 A \partial_z A - \frac{2}{r} B_\phi \partial_z B_\phi
\end{multline}

This simplifies to

\beq
J\left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_rA\right) \right) - \frac{2}{r}B_\phi \partial_z B_\phi
\eeq

Full derivation of viscous term:

\beq
\partial_z \left(\frac{1}{\reye}\left[\nabla^2 u\right]_r\right) - \partial_r \left(\frac{1}{\reye}\left[\nabla^2 u\right]_z\right)
\eeq

\beq
= \frac{1}{\reye}\left[\partial_z \left(\nabla^2 u_r - \frac{1}{r^2} u_r\right) - \partial_r \left(\nabla^2 u_z\right)\right]
\eeq

\begin{multline}
= \frac{1}{\reye} \left[ - \frac{2}{r^2} \partial_z^2 \partial_r \Psi + \frac{2}{r} \partial_z^2 \partial_r^2 \Psi + \frac{1}{r} \partial_z^4 \Psi - \frac{3}{r^4} \partial_r \Psi + \frac{3}{r^3} \partial_r^2 \Psi - \frac{2}{r^2} \partial_r^3 \Psi + \frac{1}{r} \partial_r^4 \Psi \right]
\end{multline}

\section{Recovery of Narrow Gap Equations}
\label{sec:narrow_gap_recovery}

In order to recover the narrow gap, we take the transformation from dimensional, cylindrical coordinates $(\widetilde{r},\widetilde{z})$ to dimensioneless quantities $(x,z)$. In order to do so, we use $\widetilde{r} = r_0 (1 + \delta x)$, where $\delta = (r_{out} - r_{in})/r_0 \to 0$. Then, we recall that we must change all derivatives $\partial_{\widetilde{r}} \psi = \partial_x \psi \partial_{\widetilde{r}} x$ and likewise for $\partial_{\widetilde{z}}$. In order that the $x$ and $z$ parts of the equations come in at the same order, we must choose $\widetilde{z} = r_0 \delta z$. Then, $\partial_{\widetilde{r}} x = \delta^{-1}$ and $\partial_{\widetilde{z}} z = \delta^{-1}$. When you finally expand out all terms, you should see that everything that is not $\partial^2_x$ and $\partial^2_z$ will be at higher order in $\delta$, and thus will go to zero. For example, just taking the scalar Laplacian of $B_\phi$,

\begin{equation}
  \nabla^2 B_\phi = \frac{1}{r} \frac{\partial}{\partial r}\left(r \frac{\partial \psi}{\partial r} \right) + \frac{\partial^2 \psi}{\partial z^2}
\end{equation}

gives

\begin{equation}
  \frac{1}{(1 + \delta x)} \delta^{-2} \partial_x \left[(1 + \delta x) \partial_x B_\psi \right] + \delta^{-2} \partial_z^2 B_\phi,
\end{equation}
and using the binomial expansion,

\begin{equation}
  (1 - \delta x) \delta^{-2} \left(\partial_x^2 B_\psi + \delta \partial_x B_\psi + \delta x \partial_x^2 B_\psi \right) + \delta^{-2} \partial_z^2 B_\phi.
\end{equation}

When $\delta \to 0$, this leaves

\begin{equation}
  \partial_x^2 B_\phi + \partial_z^2 B_\phi,
\end{equation}
as expected. Of course, we'd also do $B_\phi \to B_y$ at the same time.

\section{Nondimensionalization}

\subsection{Controlled gap width nondimensionalization}

The idea is to parameterize the gap width rather than simply nondimensionalizing lengths by a point in the center of the channel $r_0$, as we did below.

There are many ways to do this but one option is to define

\beq
\widetilde{r} = r_0 + \widetilde{x}
\eeq

Where tildes indicate dimensional quantities. Then for $\widetilde{x}$ to stay Cartesian, we need

\beq
\widetilde{r} - r_0 \ll r_2 - r_1 = d
\eeq

where $r_2 = r_1 = d$ is the gap width. Also define $\widetilde{x} \ll d$ and $d/r_0 << 1$ for the small gap approximation. Thus when we nondimensionalize,

\beq
r = \frac{\widetilde{r}}{r_0} = 1 + \frac{\widetilde{x}}{r_0}
\eeq

Note that we could control the gap width in terms of a small parameter $\epsilon x$ instead of a fraction $\frac{x}{r_0}$, but I believe both should work the same way.

In the example Jeff worked out before, we looked at the base velocity profile in the thin gap limit. The rotation profile is given by $\Omega(r) = \Omega_0 r^{-q}$ in dimensionless quantities, and the background flow $u = r\Omega(r)$, so in our new notation:

\beq
u = \left(1 + \frac{\widetilde{x}}{r_0}\right) \Omega_0 \left(1 + \frac{\widetilde{x}}{r_0}\right)^{-q} = \Omega_0 \left(1 + \frac{\widetilde{x}}{r_0}\right)^{1-q}  
\eeq

Taylor expanding the exp(1-q) quantity, we arrive at

\beq
u = \Omega_0 \left(1 + \frac{\widetilde{x}}{r_0}\right) - \Omega_0 q \frac{\widetilde{x}}{r_0}
\eeq

Then if we define this with respect to a rest frame $\Omega_0 \left(1 + \frac{\widetilde{x}}{r_0}\right)$ and define a dimensionless $x = \widetilde{x}/r_0$, we get the Umurhan+ base flow $u = - \Omega_0 q x$.

This makes sense to me but I'm having trouble extending it to the actual equations. This implies, for example, that the term $\frac{1}{\eta} \nabla^2 B$, which in the $\hat{\phi}$ equation is

\beq
\left[\nabla^2 B\right]_\phi = \nabla^2 B_\phi - \frac{B_\phi}{r^2},
\eeq

should become

\beq
\nabla^2 B_\phi - B_\phi \left(1 - \frac{2 \widetilde{x}}{r_0}\right)
\eeq

after substitution and Taylor expansion of the $r^{-2}$. But clearly there's no $- B_\phi \left(1 - \frac{2 \widetilde{x}}{r_0}\right)$ in the thin gap equation. What am I missing?? Is this not the right place to be substituting in the new definition of r??

\subsection{Typical nondimensionalization}
The momentum equation must be nondimensionalized. Here are the definitions of all of the dimensional components:

\begin{equation}
\begin{multlined}
\widetilde{u} = \Omega_0 r_0 \delta u \\
\widetilde{x} = \delta r_0 x \\
\widetilde{\nabla} = \frac{\nabla}{\delta r_0}\\
\widetilde{t} = \frac{t}{\Omega_0}\\ 
\widetilde{B} = B_0 B \\
\widetilde{P} = P_0 P \\
\widetilde{\rho} = \rho_0 \rho 
\end{multlined}
\end{equation}

If we define the velocity scale $v_0$ as the local sound speed scale $\left(c_{s0} \equiv \sqrt{\frac{P_0}{\rho_0}}\right)$, then $\Omega_0 r_0 \delta = \sqrt{\frac{P_0}{\rho_0}}$.

Nondimensionalizing the momentum equation and dividing by a factor of $\Omega_0^2 r_0 \delta$ yields

\beq
\partial_t \mathbf{u} + \mathbf{u}\cdot \mathbf{\nabla} \mathbf{u} = \frac{1}{4 \pi \rho} \frac{B_0^2}{\rho_0 \Omega_0^2 r_0^2 \delta^2} (\mathbf{\nabla}\times\mathbf{B})\times\mathbf{B} + \frac{1}{\reye}\nabla^2 \mathbf{u} - 2 \mathbf{\Omega} \times \mathbf{u} - \mathbf{\Omega} \times (\mathbf{\Omega} \times \mathbf{r})
\eeq

If we define

\beq
\beta \equiv \frac{P}{P_{mag}} = \frac{\rho_0 \Omega_0^2 r_0^2 8 \pi}{B_0^2}
\eeq

where $P_{mag} = \frac{B_0^2}{8 \pi}$, then the factor in front of $(\mathbf{\nabla}\times\mathbf{B})\times\mathbf{B} $ should be $\frac{2}{\beta}$. Agreed?

\section{Perturbed Equations}

We perturb the wide gap equations according to

\beq
\mathbf{B} = B_0 \hat{z} + \mathbf{B_1},
\eeq

which is the same perturbation we used in the thin-gap construction. But wait! We perturbed the thin-gap equations according to

\beq
\mathbf{u} = -q \Omega_0 r \hat{\phi} + \mathbf{u_1}
\eeq

but I don't think this is valid in the wide-gap case. To be sure, I'm going to perturb instead by

\beq
\mathbf{u} = r\Omega(r)\hat{\phi}+ \mathbf{u_1}
\eeq

where $\Omega(r) = \Omega_0 \left(\frac{r}{r_0}\right)^{-q} = \Omega_0 \left(r\right)^{-q}$ in dimensionless coordinates (but keeping $\Omega_0$ to flag a rotational term...) Agreed?

We also add the Coriolis and centrifugal terms $-2 \mathbf{\Omega} \times \mathbf{u} - \mathbf{\Omega} \times \left(\mathbf{\Omega} \times \mathbf{r}\right)$ which expand as follows:

\beq
-2\Omega(r)\zhat \times \left( \mathbf{u_1} + r\Omega(r)\phihat\right) = -2\Omega(r)u_{r}\phihat + 2\Omega(r)u_\phi\rhat + 2r\Omega(r)^2\rhat
\eeq

\beq
-\Omega(r)\zhat \times \left(\Omega(r)\zhat \times r\rhat\right) = +\Omega(r)^2 r \rhat
\eeq

The $\rhat$ terms will have $\partial_z$ applied to them, which will destroy the $3\Omega(r)^2 r \rhat$ term and so the righthand side of the $\Psi$ equation will ultimately gain only the term $2 \Omega(r) \partial_z u_\phi \rhat$. The $\phihat$ equation gains the term $-\frac{2}{r} \Omega(r) \partial_z \Psi$ on the righthand side.

The $\Psi$ equation becomes

\begin{multline}
\label{eq:Psi_perturbed}
\frac{1}{r}\partial_t \left(\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi \right) + J\left(\Psi, \frac{1}{r^2} \left( \nabla^2 \Psi - \frac{2}{r} \partial_r\Psi\right) \right) - \frac{1}{r} 2 u_\phi \partial_z u_\phi - \partial_z\left(u_\phi \Omega(r) \right)\\
= \frac{2}{\beta} J\left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_rA\right) \right) - \frac{2}{\beta} \frac{2}{r}B_\phi \partial_z B_\phi + 2 \Omega(r) \partial_z u_\phi \\
+ \frac{1}{\reye} \left[ \nabla^2 \left(\frac{1}{r} \nabla^2 \Psi\right) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4}\partial_r\Psi\right] + \frac{2}{\beta} \frac{1}{r}B_0 \partial_z \left(\nabla^2 A - \frac{2}{r} \partial_r A\right) 
\end{multline}

Note that one of the terms gained from the base state in the $(\mathbf{u}\cdot \nabla)\mathbf{u}$ term can be combined with the term gained from the Coriolis term, but here I write them separately so we can check the equation.
Equation \ref{eq:uy} becomes

\begin{multline}
  \label{eq:uy_perturbed}
\partial_t \uphi + \frac{J(\psi, \uphi)}{r} + \frac{\uphi \partial_z \psi}{r^2} + \partial_z\Psi\left( \frac{2}{r}\Omega(r) + \partial_r \Omega(r) \right) = \\
\frac{2}{\beta} \frac{J(A, B_\phi)}{r} + \frac{2}{\beta} \frac{B_\phi \partial_z A}{r^2} + \frac{1}{\reye} \left( \nabla^2 \uphi - \frac{\uphi}{r}\right) + \frac{2}{\beta} B_0 \partial_z B_\phi -\frac{2}{r} \Omega(r) \partial_z \Psi
\end{multline}

%$\frac{2}{r} \partial_z \Psi \Omega(r) + \partial_z\Psi \partial_r \Omega(r)
%=\partial_z\Psi\left( \frac{2}{r}\Omega(r) + \partial_r \Omega(r) \right)$

%\begin{equation}
 % \label{eq:uy_perturbed}
%\partial_t \uphi + \frac{J(\psi, \uphi)}{r} + \frac{\uphi \partial_z \psi}{r^2} - \frac{2}{r}q \Omega_0 \partial_z \Psi= \frac{2}{\beta} \frac{J(A, B_\phi)}{r} + \frac{2}{\beta} \frac{B_\phi \partial_z A}{r^2} + \frac{1}{\reye} \left( \nabla^2 \uphi - \frac{\uphi}{r}\right) + \frac{2}{\beta} B_0 \partial_z B_\phi
%\end{equation}

%The only term gained in the $(\hat{r}, \hat{z})$ component of the induction equation is $(B_0 \hat{z} \cdot \nabla) \mathbf{u_1}$, and

Equation \ref{eq:A} picks up $+\Omega(r) B_\phi$ from the term $- ({u_0}\hat{\phi} \cdot \nabla)\mathbf{B_1}$ and $-\Omega(r) B_\phi$ from the term $+(\mathbf{B_1} \cdot \nabla) \hat{\phi}$ and, so those cancel and the equation becomes

\beq
 \label{eq:A_perturbed}
\partial_t A = \frac{1}{r} J(A, \psi) + \frac{1}{\reym} \left( \nabla^2 A - \frac{2 \partial_r A}{r} \right) + B_0 \partial_z \Psi
\eeq

Note that this is perfectly analogous to the thin-gap version of this equation.

The $\hat{\phi}$ component of the induction equation, Equation \ref{eq:By}, becomes

\begin{multline}
  \label{eq:By_perturbed}
\partial_t B_\phi = \frac{1}{r} J(A, \uphi) + \frac{1}{r} J(B_\phi, \psi)\\
+ \frac{1}{r^2} B_\phi \partial_z \psi - \frac{1}{r^2} \uphi \partial_z A + \frac{1}{\reym} \left(\nabla^2 B_\phi - \frac{1}{r^2} B_\phi \right) + B_0 \partial_z u_\phi + \partial_z A \left( \frac{2}{r}\Omega(r) + \partial_r \Omega(r) \right) %- \frac{1}{r} q \Omega_0 \partial_z A
\end{multline}

\section{Matrix Formulation}

This is all pending a rigorous check of Equations \ref{eq:Psi_perturbed} - \ref{eq:By_perturbed}, which Jeff is working on, but I'll start putting this into a matrix construction.

First, the nonlinear vector, on the lefthand side of the equation, is

\beq\label{eq:nonlinearvector}
\mathbf{N} = \left[\begin{matrix}
J\left(\Psi, \frac{1}{r^2} \left(\nabla^2 \Psi - \frac{2}{r}\partial_r \Psi \right) \right) - \frac{2}{\beta} J \left(A, \frac{1}{r^2} \left( \nabla^2 A - \frac{2}{r} \partial_r A\right)\right) - \frac{2}{r} u_\phi \partial_z u_\phi + \frac{2}{\beta} \frac{2}{r} B_\phi \partial_z B_\phi \\
\frac{1}{r} J \left(\Psi, u_\phi\right) - \frac{1}{r} \frac{2}{\beta} J \left(A, B_\phi\right) + \frac{1}{r^2} u_\phi \partial_z \Psi - \frac{2}{\beta} \frac{1}{r^2} B_\phi \partial_z A \\
-\frac{1}{r} J \left(A, \Psi\right)\\
-\frac{1}{r} J \left(A, u_\phi\right) - \frac{1}{r} J \left(B_\phi, \Psi\right) - \frac{1}{r^2} B_\phi \partial_z \Psi + \frac{1}{r^2} u_\phi \partial_z A
\end{matrix}\right] 
\eeq

Note that these differ from the thin-gap nonlinear terms not only because of the curvature terms in the Jacobians, but also because of the additional advective terms in all but the $\mathbf{A}$ equation.

The $\partial_t$ terms are grouped together into

\beq
\partial_t D = \partial_t \left[\begin{matrix}
\frac{1}{r}\nabla^2 - \frac{2}{r^2} \partial_r & 0 & 0 & 0\\
0 & 1 & 0 & 0\\
0 & 0 & 1 & 0\\
0 & 0 & 0 & 1
\end{matrix}\right]
\eeq

For the righthand side of the equation, I'll follow the same process that I did with the thin-gap. That is I'll first separate the matrices by physical meaning, and then regroup terms based on $z$-dependence. The RHS can be described by three matrices: one representing the rotational terms, one representing the background field terms, and one representing the viscous/resistive terms.

\begin{multline}
RHS = \left[\begin{matrix}
3 \Omega(r) \partial_z u_\phi \\
-\frac{4}{r} \Omega(r)\partial_z \Psi - \partial_r \Omega(r) \partial_z \Psi \\
0 \\
\frac{2}{r} \Omega(r) \partial_z A + \partial_r \Omega(r) \partial_z A
\end{matrix}\right]
+ 
\left[\begin{matrix}
\frac{2}{\beta} \frac{1}{r} B_0 \partial_z \left(\nabla^2 A - \frac{2}{r} \partial_r A\right)\\
\frac{2}{\beta} B_0 \partial_z B_\phi \\
B_0 \partial_z \Psi\\
B_0 \partial_z u_\phi
\end{matrix}\right]
+
\left[\begin{matrix}
\frac{1}{\reye} \left[ \nabla^2 \left(\frac{1}{r} \nabla^2 \Psi\right) - \frac{1}{r^3} \partial_r^2 \Psi - \frac{1}{r^4} \partial_r \Psi\right]\\
\frac{1}{\reye} \left(\nabla^2 u_\phi - \frac{1}{r} u_\phi\right)\\
\frac{1}{\reym} \left(\nabla^2 A - \frac{2}{r} \partial_r A\right)\\
\frac{1}{\reym} \left(\nabla^2 B_\phi - \frac{1}{r^2} B_\phi\right)
\end{matrix}\right] \\
=
\left[\begin{matrix}
0 & 3\Omega(r) \partial_z & 0 & 0 \\
-\frac{4}{r} \Omega(r) \partial_z - \partial_r \Omega(r) \partial_z & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & \frac{2}{r} \Omega(r) \partial_z + \partial_r \Omega(r) \partial_z & 0
\end{matrix}\right]
+\\
\left[\begin{matrix}
0 & 0 & \frac{2}{\beta} \frac{1}{r} B_0 \partial_z \left(\nabla^2 - \frac{2}{r} \partial_r\right) & 0\\
0 & 0 & 0 & \frac{2}{\beta} B_0 \partial_z\\
B_0 \partial_z & 0 & 0 & 0 \\
0 & B_0 \partial_z & 0 & 0
\end{matrix}\right]
+ 
\left[\begin{matrix}
\frac{1}{\reye} \left(\nabla^2\left(\frac{1}{r}\nabla^2\right) - \frac{1}{r^3}\partial_r^2 - \frac{1}{r^4}\partial_r\right)\\
\frac{1}{\reye} \left(\nabla^2  - \frac{1}{r} \right)\\
\frac{1}{\reym} \left(\nabla^2 - \frac{2}{r} \partial_r \right)\\
\frac{1}{\reym} \left(\nabla^2 - \frac{1}{r^2} \right)
\end{matrix}\right]
\end{multline}

We separate the viscous/resistive matrix out in terms of $\partial_z$ dependence.

\begin{multline}
\mathrm{viscous}\, \,\mathrm{terms}  = \\
 \left[\begin{matrix}
\frac{1}{\reye} \left(-\frac{3}{r^4} \partial_r + \frac{3}{r^3} \partial_r^2 - \frac{2}{r^2} \partial_r^3 + \frac{1}{r} \partial_r^4\right) & 0 & 0 & 0\\
0 & \frac{1}{\reye} \left(\partial_r^2 + \frac{1}{r} \partial_r - \frac{1}{r} \right) & 0 & 0 \\
0 & 0 & \frac{1}{\reym} \left(\partial_r^2 - \frac{1}{r} \partial_r\right) & 0 \\
0 & 0 & 0 & \frac{1}{\reym} \left(\partial_r^2 + \frac{1}{r}\partial_r - \frac{1}{r^2}\right)
\end{matrix}\right] \\
+
\partial_z^2 \left[\begin{matrix}
\frac{1}{\reye}\left(-\frac{2}{r^2}\partial_r + \frac{2}{r}\partial_r^2\right) & 0 & 0 & 0 \\
0 & \frac{1}{\reye} & 0 & 0 \\
0 & 0 & \frac{1}{\reym} & 0 \\
0 & 0 & 0 & \frac{1}{\reym}
\end{matrix}\right] \\
+
\partial_z^4\left[\begin{matrix}
\frac{1}{\reye}\frac{1}{r} & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
\end{matrix}\right]
\end{multline}

Note that the form of each of the terms is different in the first matrix.

We also separate the terms in the $B0$ matrix by $\partial_z$ dependence:

\beq
B_0 \, \mathrm{terms} = 
\partial_z \left[\begin{matrix}
0 & 0 & \frac{2}{\beta} \frac{1}{r} B_0 \left(\partial_r^2 - \frac{1}{r}\partial_r\right) & 0 \\
0 & 0 & 0 & \frac{2}{\beta}B_0\\
B_0 & 0 & 0 & 0\\
0 & B_0 & 0 & 0
\end{matrix}\right] 
+
\partial_z^3 \left[\begin{matrix}
0 & 0 & \frac{2}{\beta} \frac{1}{r} B_0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
\end{matrix}\right]
\eeq

We replace $B_0$ with our lower magnetic field strength $(1 - \epsilon^2)$:

\beq
B_0 \, \mathrm{terms} = 
(1 - \epsilon^2) \partial_z \left[\begin{matrix}
0 & 0 & \frac{2}{\beta} \frac{1}{r} \left(\partial_r^2 - \frac{1}{r}\partial_r\right) & 0 \\
0 & 0 & 0 & \frac{2}{\beta}\\
1 & 0 & 0 & 0\\
0 & 1 & 0 & 0
\end{matrix}\right] 
+
(1 - \epsilon^2) \partial_z^3 \left[\begin{matrix}
0 & 0 & \frac{2}{\beta} \frac{1}{r} & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
\end{matrix}\right]
\eeq

We now define, in analogy to the thin-gap matrices, the following:

\beq
\mathcal{L}_0 = \left[\begin{matrix}
\frac{1}{\reye} \left(-\frac{3}{r^4} \partial_r + \frac{3}{r^3} \partial_r^2 - \frac{2}{r^2} \partial_r^3 + \frac{1}{r} \partial_r^4\right) & 0 & 0 & 0\\
0 & \frac{1}{\reye} \left(\partial_r^2 + \frac{1}{r} \partial_r - \frac{1}{r} \right) & 0 & 0 \\
0 & 0 & \frac{1}{\reym} \left(\partial_r^2 - \frac{1}{r} \partial_r\right) & 0 \\
0 & 0 & 0 & \frac{1}{\reym} \left(\partial_r^2 + \frac{1}{r}\partial_r - \frac{1}{r^2}\right)
\end{matrix}\right]
\eeq

\beq
\mathcal{L}_1 = \left[\begin{matrix}
0 & 3\Omega(r) & \frac{2}{\beta} \frac{1}{r} \left(\partial_r^2 - \frac{1}{r}\partial_r\right) & 0 \\
-\frac{4}{r} \Omega(r) - \partial_r \Omega(r) & 0 & 0 & \frac{2}{\beta} \\
1 & 0 & 0 & 0 \\
0 & 1 & \frac{2}{r} \Omega(r)  + \partial_r \Omega(r) & 0
\end{matrix}\right]
\eeq

\beq
\mathcal{L}_2 = \left[\begin{matrix}
\frac{1}{\reye}\left(-\frac{2}{r^2}\partial_r + \frac{2}{r}\partial_r^2\right) & 0 & 0 & 0 \\
0 & \frac{1}{\reye} & 0 & 0 \\
0 & 0 & \frac{1}{\reym} & 0 \\
0 & 0 & 0 & \frac{1}{\reym}
\end{matrix}\right]
\eeq

\beq
\mathcal{L}_3 = \left[\begin{matrix}
0 & 0 & \frac{2}{\beta} \frac{1}{r} & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
\end{matrix}\right]
\eeq

\beq
\mathcal{L}_4 = \left[\begin{matrix}
\frac{1}{\reye}\frac{1}{r} & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
\end{matrix}\right]
\eeq

\beq
\mathcal{G} = \left[\begin{matrix}
0 & 0 & \frac{2}{\beta} \frac{1}{r} \left(\partial_r^2 - \frac{1}{r}\partial_r\right) & 0 \\
0 & 0 & 0 & \frac{2}{\beta}\\
1 & 0 & 0 & 0\\
0 & 1 & 0 & 0
\end{matrix}\right] 
\eeq

Defining

\beq
\mathcal{L}  = \mathcal{L}_0 + \mathcal{L}_1\partial_z + \mathcal{L}_2\partial_z^2 + \mathcal{L}_3\partial_z^3 + \mathcal{L}_4\partial_z^4
\eeq

and

\beq
\mathcal{\widetilde{G}} = - \mathcal{G} \partial_z - \mathcal{L}_3 \partial_z^3, 
\eeq

we have

\beq
\mathcal{D} \partial_t \mathbf{V} + \mathbf{N} = \mathcal{L}\mathbf{V} + \epsilon^2 \mathcal{\widetilde{G}} \mathbf{V}
\eeq

which we then expand:

\beq
\begin{split}
 \partial_t & \rightarrow \epsilon^2 \partial_T \\
 \partial_z & \rightarrow \partial_z + \epsilon \partial_Z
\end{split}
\eeq

and perturb:

\beq
\mathbf{V} = \epsilon\mathbf{V}_1 + \epsilon^2 \mathbf{V}_2 + \epsilon^3 \mathbf{V}_3
\eeq

Note that as we have defined these matrices, the matrices are of course different than in the thin-gap case, and the nonlinear terms are different, but the equations for behavior at each order in $\epsilon$ are identical, that is

\beq
\mathcal{L} \mathbf{V}_1 = 0
\eeq

\beq
\mathcal{L} \mathbf{V}_2 = \mathbf{N}_2 - \mathcal{\widetilde{L}}_1 \partial_Z \mathbf{V}_1
\eeq

\beq
\mathcal{D}\partial_T\mathbf{V}_1 + \mathbf{N}_3 = \mathcal{L}\mathbf{V}_3 + \mathcal{\widetilde{L}}_1 \partial_Z \mathbf{V}_2 + \mathcal{\widetilde{L}}_2\partial_Z^2 \mathbf{V}_1 + \mathcal{\widetilde{G}}\mathbf{V}_1
\eeq

are still the correct equations to orders $\epsilon$, $\epsilon^2$, and $\epsilon^3$, respectively.

\section{Expansion of Nonlinear Terms}

Expanding the terms in Equation \ref{eq:nonlinearvector}, we find.

\beq
\mathbf{N} = \epsilon^2 \mathbf{N}_2 + \epsilon^3 \mathbf{N}_3
\eeq

\beq
\begin{split}
N_2^{\Psi} & = J\left(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_1\right) + J\left(\Psi_1, -\frac{2}{r^3}\partial_r\Psi_1\right) \\
&- \frac{2}{\beta} J \left(A_1, \frac{1}{r^2} \nabla^2 A_1\right) - \frac{2}{\beta} J\left(A_1, -\frac{2}{r^3} \partial_r A_1\right) - \frac{2}{r} u_1 \partial_z u_1 + \frac{2}{\beta} \frac{2}{r} B_1 \partial_z B_1
\end{split}
\eeq

\beq
\begin{split}
N_3^{\Psi} & = J\left(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_2\right) + J\left(\Psi_2, \frac{1}{r^2} \nabla^2\Psi_1\right) + 2 J \left(\Psi_1, \frac{1}{r^2}\partial_Z\partial_z \Psi_1\right) \\
&+ J\left(\Psi_1, -\frac{2}{r^3}\partial_r \Psi_2\right) + J\left(\Psi_2, -\frac{2}{r^3}\partial_r \Psi_1\right) + \widetilde{J}\left(\Psi_1, \frac{1}{r^2} \nabla^2 \Psi_1\right) + \widetilde{J} \left(\Psi_1, -\frac{2}{r^3}\partial_r \Psi_1\right)\\
& - \frac{2}{\beta} J\left(A_1, \frac{1}{r^2}\nabla^2 A_2\right) - \frac{2}{\beta} J\left(A_2, \frac{1}{r^2}\nabla^2 A_1\right) - \frac{4}{\beta} J\left(A_1, \frac{1}{r^2}\partial_Z\partial_z A_1\right) - \frac{2}{\beta} J\left(A_1, -\frac{2}{r^3} \partial_r A_2 \right) \\
& - \frac{2}{\beta} J\left(A_2, -\frac{2}{r^3} \partial_r A_1\right) - \frac{2}{\beta} \widetilde{J} \left(A_1, \frac{1}{r^2} \nabla^2 A_1\right) - \frac{2}{\beta} \widetilde{J} \left(A_1, -\frac{2}{r^3} \partial_r A_1\right) \\
& - \frac{2}{r} u_1 \partial_z u_2 - \frac{2}{r} u_2 \partial_z u_1 - \frac{2}{r} u_1 \partial_Z u_1 + \frac{2}{\beta}\frac{2}{r} B_1\partial_z B_2 + \frac{2}{\beta}\frac{2}{r} B_2 \partial_z B_1 + \frac{2}{\beta} \frac{2}{r} B_1 \partial_Z B_1
\end{split}
\eeq

\beq
\begin{split}
N_2^{u} = \frac{1}{r} J\left(\Psi_1, u_1\right) - \frac{1}{r} \frac{2}{\beta} J\left(A_1, B_1\right) + \frac{1}{r^2} u_1 \partial_z \Psi_1 - \frac{2}{\beta}\frac{1}{r^2} B_1 \partial_z A_1
\end{split}
\eeq

\beq
\begin{split}
N_3^u & = \frac{1}{r}J\left(\Psi_1, u_2\right) + \frac{1}{r}J\left(\Psi_2, u_1\right) + \frac{1}{r}\widetilde{J} \left(\Psi_1, u_1\right) \\
& - \frac{1}{r}\frac{2}{\beta} J\left(A_1, B_2\right) - \frac{1}{r} \frac{2}{\beta} J\left(A_2, B_1\right) - \frac{1}{r}\frac{2}{\beta}\widetilde{J}\left(A_1, B_1\right) \\
& + \frac{1}{r^2} u_1\partial_z \Psi_2 + \frac{1}{r^2} u_2 \partial_z \Psi_1 + \frac{1}{r^2} u_1 \partial_Z \Psi_1 \\
& - \frac{2}{\beta} \frac{1}{r^2} B_1 \partial_z A_2 - \frac{2}{\beta} \frac{1}{r^2} B_2 \partial_z A_1 - \frac{2}{\beta} \frac{1}{r^2} B_1 \partial_Z A_1
\end{split}
\eeq

\beq
N_2^A = -\frac{1}{r} J\left(A_1, \Psi_1\right)
\eeq

\beq
N_3^A = -\frac{1}{r} J\left(A_1, \Psi_2\right) - \frac{1}{r}J\left(A_2, \Psi_1\right) - \frac{1}{r} \widetilde{J}\left(A_1, \Psi_1\right)
\eeq

\beq
N_2^B = -\frac{1}{r} J\left(A_1, u_1\right) - \frac{1}{r} J\left(B_1, \Psi_1\right) - \frac{1}{r^2} B_1 \partial_z \Psi_1 + \frac{1}{r^2} u_1 \partial_z A_1
\eeq

\beq
\begin{split}
N_3^B & = - \frac{1}{r} J\left(A_1, u_2\right) - \frac{1}{r} J\left(A_2, u_1\right) - \frac{1}{r}\widetilde{J}\left(A_1, u_1\right) \\
& - \frac{1}{r} J\left(B_1, \Psi_2\right) - \frac{1}{r} J\left(B_2, \Psi_1\right) - \frac{1}{r} \widetilde{J} \left(B_1, u_1\right) \\
& - \frac{1}{r^2} B_1\partial_z \Psi_2 - \frac{1}{r^2} B_2 \partial_z \Psi_1 - \frac{1}{r^2} B_1 \partial_Z \Psi_1 \\
& + \frac{1}{r^2} u_1 \partial_z A_2 + \frac{1}{r^2} u_2 \partial_z A_1 + \frac{1}{r^2} u_1 \partial_Z A_1
\end{split}
\eeq

\section{Notes for paper}

-- Comment on channel mode formation w.r.t. parasitic modes?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                              APPENDIX                                                             %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix

\section{Cylindrical derivatives}
\label{sec:cylindrical_deriv}

Everything here follows \url{http://farside.ph.utexas.edu/teaching/336L/Fluidhtml/node257.html}.

For a scalar field $\psi$,

\begin{equation}
  \label{eq:del_scalar}
  \nabla \psi = \frac{\partial \psi}{\partial r} \rhat + \frac{1}{r} \frac{\partial \psi}{\partial \phi} \phihat + \frac{\partial \psi}{\partial z} \zhat,
\end{equation}

and

\begin{equation}
  \label{eq:del2_scalar}
  \nabla^2 \psi = \frac{1}{r} \frac{\partial}{\partial r}\left(r \frac{\partial \psi}{\partial r} \right) + \frac{1}{r^2} \frac{\partial^2 \psi}{\partial \phi^2} + \frac{\partial^2 \psi}{\partial z^2},
\end{equation}

However, for a \emph{vector} field $\mathbf{u}$,
\begin{equation}
  \label{eq:div}
  \mathbf{\nabla \cdot u} = \frac{1}{r} \frac{\partial (r u_r)}{\partial r} + \frac{1}{r} \frac{\partial \uphi}{\partial \phi}  + \frac{\partial u_z}{\partial z}
\end{equation}

and

\begin{equation}
  \label{eq:curl}
  \mathbf{\nabla \times u} = \left( \frac{1}{r} \frac{\partial u_z}{\partial \phi} - \frac{\partial u_\phi}{\partial z}\right)\ \rhat + \left(\frac{\partial u_r}{\partial z} - \frac{\partial u_z}{\partial r}\right)\ \phihat  + \left( \frac{1}{r} \frac{\partial (r \uphi)}{\partial r} - \frac{1}{r}\frac{\partial u_r}{\partial \phi}\right)\ \zhat.
\end{equation}

We also need the $\phi$ component of the convective derivative $\mathbf{u \cdot \nabla u}$,
\begin{equation}
  \label{eq:convective_deriv_phi}
  \left[\mathbf{u \cdot \nabla u} \right]_\phi = \mathbf{u \cdot \nabla} \uphi + \frac{ u_r \uphi}{r},
\end{equation}
and finally, the vector Laplacian,
\begin{equation}
  \label{eq:vec_lap_r}
  (\nabla^2 \mathbf{u})_r = \nabla^2 u_r - \frac{u_r}{r^2} - \frac{2}{r^2} \frac{\partial \uphi}{\partial \phi}
\end{equation}

\begin{equation}
  \label{eq:vec_lap_phi}
  (\nabla^2 \mathbf{u})_\phi = \nabla^2 u_\phi + \frac{2}{r^2} \frac{\partial u_r}{\partial \phi} - \frac{u_\phi}{r^2} 
\end{equation}

\begin{equation}
  \label{eq:vec_lap_z}
  (\nabla^2 \mathbf{u})_z = \nabla^2 u_z,
\end{equation}
where $\nabla$ on the vector components is given by equation~(\ref{eq:del_scalar}).

Note that, expanding the definition of the vector Laplacian, where the cylindrical scalar Laplacian is substituted in for $\nabla^2 u_r$ and $\nabla^2 u_z$

\end{document}